Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-05-24 manuscript/manuscript.tex
Removed old commented lines in manuscript
Changes in abstract

\begin{summary}
We present a new methodology to compute the gravitational fields generated by 
any tesseroid whose density varies continuously in depth, i.e. a forward 
gravity model in spherical coordinates for varying densities on the radial 
direction.
It's based on the numerical approximation of the gravitational fields by 
applying the Gauss-Legendre Quadrature along with a preexisting adaptive 
discretization algorithm and a new density-based discretization algorithm that 
control both the accuracy and the computation time of the forward model.
The first algorithm is regulated by a predefined scalar called distance-size 
ratio ($D$) and reduces the errors due to the distance between the tesseroid 
and the computation point.
While the second one, controlled by a delta ratio ($\delta$), decreases the 
errors introduced by the variation of the density function.
We have also obtained analytical solutions for a spherical shell with variable 
density and compared them with the results of the numerical model in case of a 
linear and an exponential density function.
These comparisons allowed to obtain default values for the distance-size and 
the delta ratio in order to guarantee the accuracy of the forward model.
The resulting optimal values of $D$ for the gravity potential, the components 
of its gradient and the Marussi tensor components are 1, 2 and 8, 
respectively.
While a $\delta=0.2$ is needed for the computation of the gravitational 
potential and its gradient components and a $\delta=0.01$ must be used for 
the Marussi tensor components in order to guarantee the accuracy even for high 
varying exponential densities.
We have finally tested the new software by modelling the Neuqu\'en sedimentary 
basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S 
latitude, assigning an exponential density variation to it and computing all 
the gravitational fields it generates.
\end{summary}

@@

%~ It worth mention that either a greater $D$ or a lower $\delta$ 
%~ increases the number of discretizations of the tesseroid.
%~ Although a greater $D$ may divide it in the latitude, longitude 
%~ and/or radial direction; a lower $\delta$ will only discretize the 
%~ original tesseroid in the radial direction.
%~ This means that a decrease in $\delta$ will produce a more efficient 
%~ computation than an increase in $D$.
%~ \todo[inline]{Should we keep this paragraph? The same is explained bellow.}

@@

%~ In order to determine the value of $\delta_p$ we should compute the 
%~ differences between the numerical model and the analytical solution for 
%~ a spherical shell with an exponential density.
%~ Thus we must obtain this analytical solution and define a tesseroid 
%~ model to numerically approximate the gravitational fields.

%~ Secondly, we proposed a thin and a thick spherical shells with 1km and 
%~ 35km of thickness, respectively, modelled by a mesh of $30^\circ \times 
%~ 30^\circ$ tesseroids.
%~ We have assigned the exponential density function proposed on equation 
%~ \ref{eq:density-exp} to each spherical shell model, choosing some values 
%~ of the constant $b$ in order to test the numerical approximation for 
%~ different density variations.

%~ Finally, we computed the gravitational potential ($V$), the 
%~ vertical component of its gradient ($g_z$) and the $g_zz$ component of 
%~ the Marussi tensor on the big grid (see Table \ref{tab:grids}) by 
%~ keeping the distance-size ratio equal to the $D$ values obtained in the 
%~ density case for its corresponding gravitational field and exploring 
%~ different values of $\delta$.
%~ We then computed the percentage difference between this numerical 
%~ results and the analytical solutions.

@@

%~ We state that this method it's well suited to approximate the 
%~ effect of tesseroids with arbitrary densities, in contrast with the ones that 
%~ involve Taylor series expansion.

@@

%~ This distance-size determinations have been carried out by comparing the 
%~ analytical solutions of the gravitational fields generated by a spherical 
%~ shell with a linear density with the gravitational fields generated by the 
%~ numerical model for different values of $D$.
%~ This comparisons were taken into account to obtain a percentage difference 
%~ between the gravitational potential, the $g_z$ component of its gradient and 
%~ the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
%~ We have set the default value of density-based discretization as the minimum

%~ The distance-size ratio determinations have been made through the comparison 
%~ of the numerical approximations with the analytical solutions for a spherical 
%~ shell with variable density, which had to be analytically obtained. These 
%~ tests have been made without applying the density-based discretization 
%~ algorithm in order to isolate the former one. From the wide range of possible 
%~ density variations, we've chosen a linear, an exponential and a discontinuous 
%~ density functions. The comparisons over the first two functions have shown 
%~ that values of $D$ equal to 1, 2 and 8 for the calculation of the gravity 
%~ potential, its gradient and the Marussi tensor components, respectively, 
%~ produce an accuracy of 0.1\% or better. On the other hand, the algorithm on 
%~ its own does not guarantee this amount of accuracy in case of highly varying 
%~ exponential functions or discontinuous densities.

%~ We have performed similar tests for the density-based discretization algorithm 
%~ comparing the numerical model with the analytical solutions for a spherical 
%~ shell with various exponential densities: from low to high varying functions. 
%~ Each comparison was made for different values of $\delta$, keeping the default 
%~ values of $D$ previously obtained. This tests have shown that a value of 
%~ $\delta = 0.2$ guarantees an accuracy of 0.1\% or better on every case, while 
%~ the same results can be obtained performing this tests with the discontinuous 
%~ density function. The default value $\delta$ is the same for every gravity 
%~ field computation.

%~ Moreover, we had performed speed comparisons between tesseroids models with 
%~ homogeneous and variable densities, showing that a high varying density 
%~ function requires more computation time than a low varying one, as expected.

>>>>

\begin{summary}
We present a new methodology to compute the gravitational fields generated by 
any tesseroid whose density varies continuously in depth, i.e. a forward 
gravity model in spherical coordinates for varying densities on the radial 
direction.
It numerical approximates these gravitational fields through the 
Gauss-Legendre Quadrature along with two discretization algorithms that 
increase its accuracy by subdividing the tesseroid into smaller ones.
The first one is a preexisting adaptive discretization algorithm that reduces 
the errors due to the distance between the tesseroid and the computation 
point, while being regulated by a predefined scalar called distance-size ratio 
($D)$.
The second one, controlled by a predefined delta ratio ($\delta$), is a new 
density-based discretization algorithm that decreases the errors introduced by 
the variation of the density function.
We have also obtained analytical solutions for a spherical shell with variable 
density and compared them with the results of the numerical model in case of a 
linear and an exponential density function.
These comparisons allowed to obtain default values for the distance-size and 
the delta ratio in order to guarantee the accuracy of the forward model.
The resulting optimal values of $D$ for the gravity potential, the components 
of its gradient and the Marussi tensor components are 1, 2 and 8, 
respectively.
While a $\delta=0.2$ is needed for the computation of the gravitational 
potential and its gradient components and a $\delta=0.01$ must be used for 
the Marussi tensor components in order to guarantee the accuracy even for high 
varying exponential densities.
We have finally tested this new methodology by modelling the Neuqu\'en 
sedimentary basin, located to the east of the Andes between 32$^\circ$S and 
40$^\circ$S latitude, assigning an exponential density variation to it and 
computing all the gravitational fields it generates.
\end{summary}

@@

@@

@@

@@
